Testing Gravity Using Type Ia Supernovae Discovered by Next Generation Wide Field Imaging Surveys

### 2019-01-26 doc/article.tex
added k_max derivative and continued on with text changes

Survey performance is examined in more detail by considering how $\sigma_\gamma$ (using RSD, peculiar velocities, and their cross-correlations) 
changes with respect to the 
survey parameters  $\Omega$, $z_{max}$, $t$, and $\sigma_M$, and also with respect to differential redshift bins
within a given survey.  
\begin{description}
\item[Solid Angle $\Omega$]
The Fisher Matrix $F$ is  proportional to the survey solid angle $\Omega$ so $\sigma_\gamma \propto \sqrt{\Omega}$.
\item[Differential Redshift Bin $z$]  Certain redshifts constrain $\gamma$ more strongly than others.  If at a given moment of a survey we had
a set of SNe~Ia from which to choose, it turns out the one with the lowest redshift would be preferred.
This is demonstrated to be the case at the end of both 2- and 10-year surveys and $z_{max}=0.2$.
Figure~\ref{dvar:fig} shows
 $|\partial \sigma_\gamma / \partial z|$, which
for both surveys monotonically decreases from $z=0.01$ out to $z=0.2$.  If we had
to sculpt the distribution (say due to limited follow-up resources), the preference would be to cut out the highest redshift bins
resulting in a decreased $z_{max}$.   The optimal redshift distribution is thus the unsculpted SN-discovery distribution truncated by $z_{max}$.
\begin{figure}
\centering
\includegraphics[width=0.47\textwidth]{src/dvardz.png}
%\includegraphics[width=0.47\textwidth]{src/dvardlnt.png}
%\includegraphics[width=0.5\textwidth]{../outcosmo/zmax_.png}
\caption{$|\partial \sigma_\gamma / \partial z|$  after two and ten years for a survy with limiting  depth $z_{max}=0.2$. 
\label{dvar:fig}}
\end{figure}
\item[Redshift Depth $z_{max}$]
Increasing the survey redshift depth increases the $\gamma$ precision.  The differential improvement in $\sigma_\gamma$ plateaus at
$z_{max} \sim 0.2$ as seen in Figure~\ref{var:fig}.
 \item[Survey duration $t$; Intrinsic Magnitude Dispersion $\sigma_M$] Increasing the survey duration to increase the number density improves constraints.  The benefit is stronger for
shallow compared to deep surveys, and the strength of the  benefit drops significantly over the course of 10 years.
These trends are seen in the left panel of Figure~\ref{dvar2:fig}, which plots $|\partial \sigma_\gamma / \partial \ln{t}|$ as a function of
 $z_{max}$ for two- and ten-year surveys, and
reflect those $k$-modes that do not benefit more data because they are sample-variance limited, either because they are at low $k$ for shallow surveys
or have already been saturated over the course of the survey.
 \begin{figure}
\centering
\includegraphics[width=0.47\textwidth]{src/dvardlnt.png}
\includegraphics[width=0.47\textwidth]{src/dvardsigM.png}
%\includegraphics[width=0.5\textwidth]{../outcosmo/zmax_.png}
\caption{Left: $|\partial \sigma_\gamma / \partial \ln{t}|$  and Right: $\partial \sigma_\gamma / \partial \sigma_M$ as a function of
 $z_{max}$ for two- and ten-year surveys.
\label{dvar2:fig}}
\end{figure}
Like survey duration, intrinsic magnitude dispersion is related to survey performance through the shot noise and thus
has similar relationships with $\sigma_\gamma$.
This is evident in the plot of $\partial \sigma_\gamma/\partial \sigma_M$ shown in the right panel of Figure~\ref{dvar2:fig}.
The improvement between $\sigma_M=0.08$ and 0.15~mag dispersions is equivalent to a factor of 3.52
 in number density, or equivalently in survey duration. 
 \end{description}

From the above calculations, we find empirically that shot noise is important since higher number densities and lower intrinsic magnitude dispersion
significantly reduce $\gamma$ uncertainty, but that sample variance is also relevant  by the end of a ten-year survey.
Survey design is aided by understanding the circumstances in which  the surveys become sample variance limited.
The  rest-frame volumetric SN~Ia rate is $7.84\times 10^{-5} h^3$\,Mpc$^{-3}$\,yr$^{-1}$;
after a few years the RSD measurement is sample variance limited.  The  contributions of $P_{vv}$ and
$\sigma^2/n$ at $z=0.1$ for $\sigma_M=0.08, 0.15$~mag with number densities from 2- and 10-year surveys
can be compared using  Figure~\ref{power:fig}.  For all of these survey scenarios, sample variance and shot-noise equality 
occurs at a $k$  within the range that carries the most signal.  However, the ten-year surveys do reach the sample variance
limit if   $k_{max} < 0.2$~$h$\,Mpc$^{-1}$.

%
%
%The current sample of SNe~Ia has a low number density compared to Fundamental Plane and Tully-Fisher galaxies.
%Nevertheless, their low intrinsic-magnitude uncertainties 
%can provide peculiar velocities (expressed equivalently as peculiar magnitudes)
%of their host galaxies \citep{2006PhRvD..73l3526H,2011ApJ...741...67D}. 
%
%\section{Peculiar Velocity Surveys are Sensitive to the Growth of Structure $\lowercase{f}D$}
%In a peculiar-velocity survey of SN~Ia host galaxies, each galaxy has an observed redshift  and a SN-Ia-based distance.
%The distance maps onto a cosmological redshift given a model for the background expansion $H_0$-free.
%The  (radial) peculiar velocity is obtained from the difference due to the Doppler effect between the observed and cosmological redshifts.
%The statistical properties of the peculiar velocities are summarized through their spatial covariance or  their power spectrum $P_{vv}$.
%
%It is standard practice to express peculiar velocities using the formalism created to describe mass overdensities.
%In linear theory, the overdensity field can be expressed as
%$
%\delta (\mathbf{x},a)  = D(a) \delta (\mathbf{x})$,
%where $D(a)$ is  the linear growth factor.
%The velocity field is related to the density field through the continuity equation as $v \propto fD\delta$,
%where $f \equiv \frac{d\ln{D}}{d\ln{a}}$ is the linear growth rate.  The velocity power spectrum $P_{vv} \propto (fD)^2 P_{\delta \delta}$.
%Peculiar velocity surveys, combined with a precision measurement of the power spectrum $P_{\delta \delta}$ 
%from the CMB, provide a measure of $fD$
%(or  equivalently $f\sigma_8$\footnote{
%The parameter $\sigma_8$, the  standard deviation of overdensities in 8$h^{-1}$Mpc spheres, is 
%commonly used in place of $D$ to normalize the
%overall amplitude of  overdensities.}).
%
%
%The survey's   peculiar velocity signal  can be described by its power spectrum $P_{vv}$.  
%Non-GR models can also predict differences in the shape and the redshift-dependence of GR prediction $P_{vv}$,
%though this detail is not considered here.
%The top panel of Figure~\ref{power:fig} shows the volume-weighted power spectrum $k^3 P_{vv}$
%at $z=0$ expected from General Relativity and linear theory. Ranges of $k$ accessed by surveys with different
%maximum redshift $z_{max}$ are also indicated, where $k_{min} \approx (2 z_{max})^{-1}$  and 
%$k_{max} \approx 0.1$~$h$\,Mpc$^{-1}$ represents  the smallest
%scale where  theory predictions are reliable.
%Figure~\ref{power:fig} shows  that the bulk of the peculiar velocity signal comes from $0.02 <k< 0.1$~$h$\,Mpc$^{-1}$ modes.
%
%The probative power of a specific peculiar-velocity tracer primarily depends on its number density and the precision to which its absolute magnitudes are known.
%The current generation of peculiar velocity studies uses $10^3-10^5$ galaxies with Fundamental Plane and Tully-Fisher distances \citep{2008AJ....135.1738M, 2014MNRAS.445.2677S,
%2016AJ....152...50T}.  These galaxies have absolute magnitude uncertainties of $\sim 0.4$~mag. 
%
%
%The current sample of SNe~Ia has a low number density compared to Fundamental Plane and Tully-Fisher galaxies.
%Nevertheless, their low intrinsic-magnitude uncertainties 
%can provide peculiar velocities (expressed equivalently as peculiar magnitudes)
%of their host galaxies \citep{2006PhRvD..73l3526H,2011ApJ...741...67D}.  Existing SN~Ia samples
%have been used to test and ultimately find spatial correlations in peculiar velocities that may be attributed to the growth of structure
%\citep{2008MNRAS.389L..47A,2014MNRAS.444.3926J,2015JCAP...12..033H, 2017JCAP...05..015H}.
%However, the signal-to-noise is currently insufficient to perform a meaningful test of GR.
%

%The results are summarized in Table~\ref{tab:howlett}, where a 5\% distance (0.1~mag) uncertainty is assumed for each supernova.
%Their Figure~4b) shows how well these measurements can distinguish between $\gamma=0.42, 0.55, 0.68$
%\citep[$f(R)$, GR,  and DGP gravities,][]{2007APh....28..481L} gravities respectively.
%SNe~Ia alone can provide a $2\%$ measurement of $f\sigma_8$ at $z<0.3$ redshifts lower than where galaxy and cluster (through the
%kinematic S-Z effect)
%RSD measurements are sensitive.  A joint galaxy RSD and SN peculiar velocity will be even more powerful.
%SNe~Ia peculiar velocities will measure $f\sigma_8$ at $z<0.3$  as well as galaxy surveys will at $z \sim 0.6$ using RSD. 
%
%%{\bf Add a SN PV + DESI-like BGS column and/or a PV-only column with Cullan's code? }
%% Requires the booktabs if the memoir class is not being used
%\begin{table}
%   \centering
%   %\topcaption{Table captions are better up top} % requires the topcapt package
%   \begin{tabular}{@{} lccc @{}} % Column formatting, @{} suppresses leading/trailing space
%	\hline
%	& RSD & RSD + PV & RSD+PV\\
%	Redshift 	 & ($z$-bin) & ($z$-bin)& (cumulative)\\  \hline
%      $0.00<z<0.05$   & 66.3 & 13.9 & 13.9\\
%     $0.05<z<0.10$            & 24.6     &  7.3 & 6.5\\
%     $0.10<z<0.15$      & 14.8  & 5.8 & 4.3\\
%     $0.15<z<0.20$      & 10.6  & 5.0 & 3.3\\
%      $0.20<z<0.25$     & 8.3  & 4.4 & 2.6\\
%     $0.25<z<0.30$  & 6.8  &  4.0 & 2.2\\
%      \hline
%   \end{tabular}
%   \caption{Projected percent uncertainties in $f\sigma_8$ from a 10-year LSST SN survey with 5\% distance uncertainties from
%   \citet{2017ApJ...847..128H}. RSD is from clustering, and RSD~+~PV is from joint clustering and peculiar velocities.
%   The cumulative column shows the effective uncertainty from combining its and all other shallower redshift bins.}
%   \label{tab:howlett}
%\end{table}
%
%Baryonic structures are sensitive to $fD$ through redshift space distortions (RSD), 
%which on large scales gives the combination $(b + f \mu^2)D$ where $b$ is the bias between the tracer and dark matter
%and $\mu$ gives the angular separation of Fourier modes to the line-of-sight. 
%The precision in  $fD$ derived from the baseline LSST WFD ten-year SN~Ia discoveries has been projected by \citet{2017ApJ...847..128H},
%from RSD of supernovae and peculiar velocities.
%

%Baryonic structures and peculiar velocities within the same volume are induced by the same overdensities  making their cross-correlation 
%insensitive to sample variance \citep{2007PhRvL..99h1301G}: galaxy and peculiar velocity surveys provide synergistic constraints more powerful
%than their naive sum.

%\section{Probing Gravity}
%The growth of structure depends on the expansion history of the Universe, the nature and density of its contents, and gravity. 
%It is therefore a powerful probe of cosmology and dark energy.  Growth of structure can be measured from the light-emittting structures
%in the Universe and from the peculiar velocities of test masses therein.
%Peculiar velocities are the motions, on top of the cosmological expansion, caused by the gravitational attraction
%and repulsion of density inhomogeneities in the Universe.  The peculiar velocity of an object with a known absolute magnitude
%can be determined from its observer-frame magnitude and redshift. For a given background cosmology, the observed magnitude provides
%an estimate of the cosmological redshift, the peculiar velocity is then the difference between the cosmological and CMB-frame redshifts.

%Correlations between peculiar velocities are sensitive to  $f\sigma_8 (H_0 d_L)^{-1}$.
%Although the velocities may be measured off of biased tracers, the tracers' dynamics are driven by all mass (including dark matter) and so are bias-free.
%As
%mentioned earlier, peculiar velocities are measured relative to the background cosmological expansion that
%leads to the dependence on $H_0 d_L$. (Note that the combination $H_0 d_L$ is independent of $H_0$.)

%
%The projected precisions for LSST-discovered SNe~Ia have a number of interesting features. 
%Despite the significant gain in volume and numbers of supernovae, the $f\sigma_8$ uncertainty in increasing redshift bins asymptotes such that 
%there is little benefit in going beyond
%$z=0.3$.
%At $z<0.2$ the constraining power on  $f\sigma_8$ 
%comes primarily from peculiar velocities
%but by $z=0.3$ RSD is more important.

%
%\section{Designing a Supernova Survey to Measure $\MakeLowercase{f}\sigma_8$}
%\subsection{Fisher Matrix of Velocity-Galaxy Surveys}
%
%
%
%\citet{2017ApJ...847..128H, 2017MNRAS.464.2517H} project the precision in $f\sigma_8$, summarized in   Table~\ref{tab:howlett},
% obtained by  complementary galaxy and peculiar velocity surveys in a common volume.
%Their analysis is performed in Fourier space, 
%where the Fisher information matrix of parameters $\lambda_i$ from random Gaussian field with mean zero is given by
%\begin{align}
%F_{ij} & = \frac{V}{2}\int \frac{d^3k}{(2\pi)^3} \text{Tr}\left[ C^{-1} \frac{\partial C}{\partial \lambda_i} C^{-1}
%\frac{\partial C}{\partial \lambda_j} \right]\\
%& = \frac{\Omega}{8\pi^2} \int_0^{r_{max}}  \int_{k_{min}}^{k_{max}}  \int_{-1}^{1} r^2 k^2 \text{Tr}\left[ C^{-1} \frac{\partial C}{\partial \lambda_i} C^{-1}
%\frac{\partial C}{\partial \lambda_j} \right] d\mu\,dk\,dr
%\label{fisher:eqn}
%\end{align}
% and covariance
%\begin{equation}
%C(k,\mu)  =
%  \begin{bmatrix}
%   P_{gg}(k,\mu) + \frac{1}{n_g} &
%   P_{vg}(k,\mu)  \\
%   P_{vg}(k,\mu)  &
%  P_{vv}(k,\mu) + \frac{\sigma^2}{n}
%   \end{bmatrix}.
%\label{cov:eq}
%\end{equation}
%
%
%The covariance depends on $P_{vv}$, the galaxy power spectrum $P_{gg}\propto (b\sigma_8 + f\sigma_8\mu^2)^2$, and the galaxy-velocity cross-correlation $P_{vg}
%\propto  (b\sigma_8 + f\sigma_8\mu^2)f\sigma_8$, where $b$ is the galaxy bias.
%The galaxy number density is $n_g$, the number density of velocity probes is $n$, and the per-probe velocity dispersion
%is $\sigma$. 
%The integrand in the Fisher-element calculation
%\begin{equation}
%I = \text{Tr}\left[ C^{-1} \frac{\partial C}{\partial (f\sigma_8)} C^{-1}
%\frac{\partial C}{\partial  (f\sigma_8)} \right]
%\end{equation}
%represents the contribution to  $F_{f\sigma_{8}\,f\sigma_{8}}$ from a
%differential Fourier
%volume element.  
%
%%
%%\begin{figure}
%%\centering
%%\includegraphics[width=0.5\textwidth]{src/new.png}
%%%\includegraphics[width=0.5\textwidth]{../outcosmo/zmax_.png}
%%\caption{ $I/I_{max}$ for $\mu=1, 0.5$  where $I=\text{Tr}\left(C^{-1}C_{,f\sigma_8}C^{-1}C_{,f\sigma_8}\right)$ represents the contribution of a single mode to
%%the Fisher element $F_{f\sigma_8 f\sigma_8}$ and $I_{max}$ is the sample variance limit.  The dashed black curve represents galaxy and velocity surveys in
%%independent volumes.  Two- and ten-year SN number densities are given by dotted and solid color lines respectively.  Different colors represent varying levels
%%of intrinsic velocity dispersion.  The bottom solid horizontal lines show the approximate range of $k$ expected to be used in surveys with corresponding
%%redshift depths $z_{max}$.
%%\label{shot:fig}}
%%\end{figure}
%
%
%\subsection{Precision in $f\sigma_8$ and Survey Design Parameters}
%First, we demonstrate that the realistic surveys being considered have significant shot noise contributions in their error budgets. 
%
%The $I/I_{max}$ of these  scenarios are shown in the bottom two panels of Figure~\ref{power:fig}.
%Generally, $k<0.01$ modes have $I/I_{max}\sim 1$ and so are sample-noise limited.
%Individual $k>0.01$ modes have  $I/I_{max}<1$ meaning  that shot noise
%is significant.  Recall that collectively, $k>0.01$ modes are the important contributors to the peculiar-velocity signal.
%Basically all  supernovae found in  the next decade contribute to improving the statistical measurement of $f\sigma_8$.
%The figure also shows how the statistical gain from additional supernovae diminish as the survey duration increases, and when they are at lower-redshift  (by virtue of their lower intrinsic velocity dispersion).
%\subsection{Survey Parameters that Determine Scientific Performance}

%The survey parameters that  enter the Fisher matrix are the volume $V$, which
%in turn can be parameterized by the survey solid angle $\Omega$ and the redshift depth $z_{max}$ ; the number density of sources $n$; and the intrinsic
%magnitude dispersion $\sigma_M$, where magnitude and velocity dispersions are related at low-redshift by $\sigma_{M} \approx \frac{5}{\ln{10}} \frac{1+z}{z} \sigma$.

%
%
%\begin{itemize}
%\item Solid Angle $\Omega$:
%The variance in $f\sigma_8$ is inversely proportional to solid angle $\Omega$.  The baseline
%LSST survey covers 18,000 sq.~deg, at $-75 \lesssim \delta \lesssim15$ avoiding the Galactic plane.  Larger solid-angle coverage for $z<0.3$ SN~Ia searches, beyond
%the LSST baseline, benefits peculiar-velocity science.  Complementary northern-hemisphere surveys and LSST-expanded or independent
%coverage of the southern equatorial pole, could significantly increase
%the sky coverage with a corresponding decrease in the variance of $f\sigma_8$.
%\item Redshift Depth $z_{max}$:
%Increasing the survey redshift depth increases the $f \sigma_8$ precision, but when resource-limited is it worth doing so at the expense of neglecting lower redshift
%SNe? This is answered in  Table~\ref{tab:howlett}, which shows that after 10 years of LSST the lowest redshifts have the
%stronger probative power and benefit most from peculiar velocities.  Although
%a differential increase in redshift depth increases the volume in Eq,~\ref{fisher:eqn}, the degradation in  intrinsic peculiar velocity $\propto z/(1+z)$
%and the fact that shot noise is still important leads to the conclusion that a low-redshift tracer is more valuable than one at higher redshift.
%
%%For the number densities generated by LSST and the SN~Ia intrinsic magnitude dispersion, the noise is not sample-variance
%%limited.  We have run simulations of LSST SN surveys
%%and calculated the signal-to-noise (STON) of $f\sigma_8$ from the peculiar-velocity correlations. These simulations show, as seen in the left plot
%%of  Figure~\ref{notshot:fig}, that an increase in the number density of tracers leads to a significant improvement in STON.
%%It is also worth noting that sample variance is not negligible, as seen by the non-linear dependence of the  STON on $\sigma_M^2/n$. 
%%
%%\begin{figure}
%%\centering
%%\includegraphics[width=0.48\textwidth]{../outcosmo/notshot.png}
%%\includegraphics[width=0.48\textwidth]{../outcosmo/perSN.png}
%%\caption{Left: STON of $f\sigma_8$ from simulated LSST SN peculiar-velocities only (no RSD).  There is a small level of scatter in the plots idue to the random
%%realizations made in simulating the surveys.
%%For several limiting redshift depths, the STON is plotted as a
%%function of $\sigma_M^2/n$, normalized to intrinsic magnitude dispersion $\sigma_M=0.1$~mag and 10-year LSST number densities.
%%Right: The effective contribution to the STON per supernova, quantified as STON divided by the root of the number of tracers.
%%\label{notshot:fig}}
%%\end{figure}
%%
%%The effective contribution to the STON per supernova, quantified as STON divided by the root of the number of tracers, is higher for lower
%%$z_{max}$.  This is shown in the
%%right plot of   Figure~\ref{notshot:fig}.
%%
%
%
%\item Intrinsic Magnitude Dispersion $\sigma_M$:
%Even after 10-years, a SN peculiar velocity survey still has significant shot noise.
%A decrease in the supernova intrinsic magnitude dispersion leads to a significant improvement in  the STON of $f\sigma_8$.
%%as shown in Figure~\ref{notshot:fig} in parallel to our earlier discussion of the dependence of $n$.  We are not sample-variance
%%limited so the $\sigma^2_M/n$-term
%%in the Fisher matrix is important.
%Survey follow-up strategy,
%which determines which data are available per SN~Ia, influences the accuracy to which their distances can be determined. 
%This dependence is non-trivial: the improvement between $\sigma_M=0.08$ and 0.15~mag dispersions is equivalent to a factor of 3.52
% in number density, or equivalently in survey duration. 
%\end{itemize}

\subsection{Designing an SN~Ia Peculiar Velocity Survey}
There are different options to consider in the design of a SN~Ia peculiar velocity survey, though at this point it is difficult to assess which is optimal
or most cost effective.  While there are a variety
of recently identified indicators of SN~Ia diversity that improve upon the 2-parameter model (SALT2) currently used to determine distances, 
there is not yet an umbrella model that simultaneously captures all indicators, their correlations, and characterizes the model residuals.
Here we  make several general conclusions
based upon the above findings, while noting that 
constructing such a model is a parallel
endeavor that will benefit peculiar-velocity  and other SN~Ia science.

@@

\begin{itemize}
\item Spectroscopic transient classification:  SNe~Ia are defined through the absence of Hydrogen and the presence of the 6150~\AA~Silicon P-Cygni feature.
The intrinsic magnitude dispersion depends on the purity of the sample. Conservatively, we consider
a pure SN~Ia sample obtained through
spectroscopic classification; the requisite follow-up resources needed for complete $z<0.3$ coverage is conceivable. 
The feasibility of using photometric classification is an important subject of research
that can have implications for survey planning, particularly in extending the redshift range.
\item Spectroscopic redshifts:  Peculiar velocities are extracted directly from redshift measurements.  Redshift uncertainties of $>0.5$\% contribute significantly
to the error budget.  We thus conclude that there is the need for spectroscopic $R>200$ host-galaxy redshifts.  In most cases this redshift will be
available in the classification spectrum.  Planned bright-galaxy redshift surveys can also observe a significant fraction of nearby SN hosts.
\item Role of LSST:
LSST can be taken to be a SN~Ia discovery and distance machine, or a discovery machine only.   LSST is expected to discover all $z<0.3$ SNe~Ia before maximum light in its active Wide Fast Deep (WFD) survey area.
The nominal cadence  
 produces sparse per-band light curves, which limits the intrinsic magnitude dispersion possible with LSST data alone to $\sigma_M \approx 0.15$~mag.
The wide-field of LSST is
 well suited for discoveries over a larger area of sky, but does not provide significant multiplex advantage for the low surface-density of active
$z<0.3$ SNe~Ia.  LSST is suited for maximizing $\Omega$ while other (cheaper) resources that supplement LSST photometry can improve $\sigma_M$.
The ultimate WFD observing strategy determines how well $\sigma_8$ can be determined with LSST data alone, and the amount of external resources needed  (if any) to
reach a targeted $\sigma_M$.

@@

\item Follow-up resources to determine per-SN distances:
Supplemental non-LSST follow-up data in the form of improved temporal light-curve sampling (accurate rise and decline times), 
expanded (UV, NIR) wavelength coverage, and spectral features
can access high-fidelity SN~Ia models with lower intrinsic magnitude dispersion and residual systematic bias.
For example, infrared data \citep{2012MNRAS.425.1007B} or spectrophotometry at peak brightness
 \citep{2015ApJ...815...58F} are projected to give $\sigma_M \lesssim 0.08$~mag.

@@

Modeling $k_{max}$ is a high as possible.

\section{Conclusions}
SNe~Ia are already powerful probes of the homogeneous cosmological expansion of the Universe.  In the next decade.
high-cadence, wide-field imaging surveys, together with improved precision in their distance determinations, will make SNe~Ia powerful probes
of the gravity induced-motion caused by the inhomogeneous Universe.  SNe~Ia peculiar velocities at $z<0.3$ will measure $f \sigma_8$ significantly better
than galaxy surveys, at lower redshifts that provide better leverage to test gravity models.
While imaging surveys will provide a steady stream of SNe~Ia, a coordinated plan of follow-up is required to take advantage of
their probative power.  The resources necessary to follow hundreds of thousands of SNe depend on specific follow-up choices, and 
access to those resources will define the redshift limits of the survey.  Fortunately, the lowest-redshift, and hence brightest, supernovae
are of the highest interest, so that a modest suite of $\sim 5$ 2--4m telescopes should be capable of providing leading measurements of
$f\sigma_8$ at $z\sim 0.1$.

>>>>

Survey performance is examined in more detail by considering how $\sigma_\gamma$ (using RSD, peculiar velocities, and their cross-correlations) 
changes with respect to the 
survey parameters  $\Omega$, $z_{max}$, $t$, and $\sigma_M$, and also with respect to differential redshift bins
within a given survey.  Though not directly a survey parameter, we also examine changes with respect to $k_{max}$.
\begin{description}
\item[Solid Angle $\Omega$]
The Fisher Matrix $F$ is  proportional to the survey solid angle $\Omega$ so $\sigma_\gamma \propto \sqrt{\Omega}$.
\item[Differential Redshift Bin $z$]  Certain redshifts constrain $\gamma$ more strongly than others.  If at a given moment of a survey we had
a set of SNe~Ia from which to choose, it turns out the one with the lowest redshift would be preferred.
This is demonstrated to be the case at the end of both 2- and 10-year surveys and $z_{max}=0.2$.
The top panel of Figure~\ref{dvar:fig} shows
 $|\partial \sigma_\gamma / \partial z|$, which
for both surveys monotonically decreases from $z=0.01$ out to $z=0.2$.  If we had
to sculpt the distribution (say due to limited follow-up resources), the preference would be to cut out the highest redshift bins
resulting in a decreased $z_{max}$.   The optimal redshift distribution is thus the unsculpted SN-discovery distribution truncated by $z_{max}$.
\begin{figure}
\centering
\includegraphics[width=0.4\textwidth]{src/dvardz.png}\\
\includegraphics[width=0.4\textwidth]{src/dvardlnt.png}
\includegraphics[width=0.4\textwidth]{src/dvardsigM.png}
%\includegraphics[width=0.4\textwidth]{src/dvardkmax.png}
%\includegraphics[width=0.47\textwidth]{src/dvardlnt.png}
%\includegraphics[width=0.5\textwidth]{../outcosmo/zmax_.png}
\caption{Top: $|\partial \sigma_\gamma / \partial z|$  after two and ten years for a survy with limiting  depth $z_{max}=0.2$. 
Bottom Left: $\sigma_\gamma^{-1}|\partial \sigma_\gamma / \partial \ln{t}|$; Bottom Right: $\sigma_\gamma^{-1}\partial \sigma_\gamma / \partial \sigma_M$each as a function of
 $z_{max}$ for two- and ten-year surveys.
\label{dvar:fig}}
\end{figure}
\item[Redshift Depth $z_{max}$]
Increasing the survey redshift depth increases the $\gamma$ precision.  The differential improvement in $\sigma_\gamma$ plateaus at
$z_{max} \sim 0.2$ as seen in Figure~\ref{var:fig}.
 \item[Survey duration $t$; Intrinsic Magnitude Dispersion $\sigma_M$] 
Increasing the survey duration  increases the number density of discovered supernovae. As
shot-noise is relevant for all the surveys considered, the increased survey duration improves the
precision in $\gamma$. The surveys that benefit most are those more susceptible to shot noise:
higher redshifts  due to their higher velocity dispersion and shorter surveys  due to their lower number 
densities.
These trends are shown in the bottom-left panel of Figure~\ref{dvar:fig}, which plots $\sigma_\gamma^{-1}|\partial \sigma_\gamma / \partial \ln{t}|$ as a function of
 $z_{max}$ for two- and ten-year surveys.
Like survey duration, intrinsic magnitude dispersion is related to survey performance through the shot noise and thus
has a similar relationship with $\sigma_\gamma$.
This is evident in the plot of $\sigma_\gamma^{-1}\partial \sigma_\gamma/\partial \sigma_M$ shown in the bottom-right panel of Figure~\ref{dvar:fig}.
 \item[Minimum length scale, maximum wavenumber $k_{max}$]  There is a minimum length scale at which density and velocity distributions are reliably predicted from theory and
 are Gaussian-distributed.  Changes in this scale engender changes in the $\gamma$ precision, as    $\sigma_\gamma^{-1}\partial \sigma_\gamma/\partial k_{max} = 0.0050$ at $k_{max}=0.1h$\,Mpc$^{-1}$
and is survey-independent.
 \end{description}

From the above calculations, we find empirically that shot noise is important
but that sample variance is non-negligible at the end of longer surveys or when the intrinsic magnitude dispersion is small.
Fine tuning
survey design is aided by understanding when surveys become sample variance limited.
After a few years the RSD measurement is sample variance limited,
the  rest-frame volumetric SN~Ia rate is $7.84\times 10^{-5} h^3$\,Mpc$^{-3}$\,yr$^{-1}$  
For peculiar velocities,
the  contributions of $P_{vv}$ and
$\sigma^2/n$ at $z=0.1$ for $\sigma_M=0.08, 0.15$~mag with number densities from 2- and 10-year surveys
can be compared using  Figure~\ref{power:fig}.  For these survey scenarios, sample variance and shot-noise equality 
occurs at a $k$  within the range that carries the most signal; both sample variance and shot-noise limited modes contribute to the constraints.
If $k_{max}$ were to decrease moderately some of the surveys could become sample variance limited, whereas increases in $k_{max}$ introduce additional
 shot-limited modes.

There are different options to consider in the design of a SN~Ia peculiar velocity survey, though at this point it is difficult to assess which is optimal
or most cost effective.  While there are a variety
of recently identified indicators of SN~Ia diversity that improve upon the 2-parameter model (SALT2) currently used to determine distances, 
there is not yet an umbrella model that simultaneously captures all indicators, their correlations, and characterizes the model residuals.
Here we  make several general conclusions
based upon the above findings, while noting that 
constructing such a model is a parallel
endeavor that will benefit peculiar-velocity  and other SN~Ia science.

@@

\begin{itemize}
\item Spectroscopic transient classification:  SNe~Ia are defined through the absence of Hydrogen and the presence of the 6150~\AA~Silicon P-Cygni feature.
The intrinsic magnitude dispersion depends on the purity of the sample. Conservatively, we consider
a pure SN~Ia sample obtained through
spectroscopic classification; the requisite follow-up resources needed for complete $z<0.3$ coverage is conceivable. 
The feasibility of using photometric classification is an important subject of research
that can have implications for survey planning, particularly in extending the redshift range.
\item Spectroscopic redshifts:  Peculiar velocities are extracted directly from redshift measurements.  Redshift uncertainties of $>0.5$\% contribute significantly
to the error budget.  We thus conclude that there is the need for spectroscopic $R>200$ host-galaxy redshifts.  In most cases this redshift will be
available in the classification spectrum.  Planned bright-galaxy redshift surveys can also observe a significant fraction of nearby SN hosts.
\item Role of the rolling search (e.g.\ LSST):
LSST can be taken to be a SN~Ia discovery and distance machine, or a discovery machine only.   LSST is expected to discover all $z<0.3$ SNe~Ia before maximum light in its active Wide Fast Deep (WFD) survey area.
The nominal cadence  
 produces sparse per-band light curves, which limits the intrinsic magnitude dispersion possible with LSST data alone to $\sigma_M \approx 0.15$~mag.
The wide-field of LSST is
 well suited for discoveries over a larger area of sky, but does not provide significant multiplex advantage for the low surface-density of active
$z<0.3$ SNe~Ia.  LSST is suited for maximizing $\Omega$ while other (cheaper) resources that supplement LSST photometry can improve $\sigma_M$.
The ultimate WFD observing strategy determines how well $\sigma_8$ can be determined with LSST data alone, and the amount of external resources needed  (if any) to
reach a targeted $\sigma_M$.

@@

\item Follow-up resources to determine per-SN distances:
Supplemental non-LSST follow-up data in the form of improved temporal light-curve sampling (accurate rise and decline times), 
expanded (UV, NIR) wavelength coverage, and spectral features
can access high-fidelity SN~Ia models with lower intrinsic magnitude dispersion and residual systematic bias.
For example, infrared data \citep{2012MNRAS.425.1007B} or spectrophotometry at peak brightness
 \citep{2015ApJ...815...58F} are projected to give $\sigma_M \lesssim 0.08$~mag.  The improvement between $\sigma_M=0.08$ and 0.15~mag dispersions is equivalent to a factor of 3.52
 in number density, or equivalently in survey duration.

@@

SNe~Ia are already powerful probes of the homogeneous cosmological expansion of the Universe.  In the next decade.
high-cadence, wide-field imaging surveys, together with improved precision in their distance determinations, will make SNe~Ia powerful probes
of the gravity induced-motion caused by the inhomogeneous Universe.  SNe~Ia peculiar velocities at $z<0.3$ will measure $f \sigma_8$ significantly better
than galaxy surveys, at lower redshifts that provide better leverage to test gravity models.
While imaging surveys will provide a steady stream of SNe~Ia, a coordinated plan of follow-up is required to take advantage of
their probative power.  The resources necessary to follow hundreds of thousands of SNe depend on specific follow-up choices, and 
access to those resources will define the redshift limits of the survey.  Fortunately, the lowest-redshift, and hence brightest, supernovae
are of the highest interest, so that a modest suite of $\sim 5$ 2--4m telescopes should be capable of providing leading measurements of
$f\sigma_8$ at $z\sim 0.1$.
